### 2013-03-28 doc/sosp13/evaluation.tex
data in table form
evaluation.tex

% \begin{figure}[ht]
%     \begin{center}
%       \includegraphics[width=0.5\textwidth]{results/benchmarks.pdf}
%     \end{center}
%     \caption{Comparing \Grappa with XMT, hand-tuned MPI and UPC.}
%     \label{fig:grappa-comparisons}
% \end{figure}

\begin{table}[htb]
\begin{center}
\begin{tabular}{l|c|c|c|c}
&	Cora &      XMT	&        MPI	&   UPC \\ \hline
GUPS	& 1	& 2.23	& 0.11& 	-- \\ 
BFS	& 1 &	4.33	& 14.37 & 	-- \\ 
Intsort	& 1	& 3.77 &	5.88	& -- \\
UTS (T1)	& 1	& 0.38	& -- &	0.09 \\ 
Pagerank	& 1	& 4.35	& 4.87 &	-- \\ 
\end{tabular}
\end{center}
\caption{Comparing \Grappa with XMT, hand-tuned MPI and UPC. Numbers are presented as throughput normalized to \Grappa.}
\label{tab:grappa-comparisons}
\end{table}

Table~\ref{tab:grappa-comparisons} shows the results. Overall, \Grappa is
within striking distance of XMT and well beyond UPC performance (UTS only).
However, it underperforms compared to hand-tuned MPI. Nevertheless, this
comparison needs to be taken with a grain of salt for several reasons. First,
we have not spent as much time tuning \Grappa's implementations. This is
supported by inspecting the BFS MPI implementation we used, which clearly
shows that it employs a lot of algorithmic-specific optimizations. In fact,
some of these optimizations resemble some of what \Grappa does automatically,
like message aggregation. Second, the GUPS results presented earlier suggests
that \Grappa performance can do a lot better.

@@

To characterize the effectiveness of aggregation, we collected
histograms of aggregated message sizes and application memory request
latency on BFS and UTS.

Figure \ref{fig:grappa-message-size} shows the distribution of message
sizes we are able to aggregate before aggregation times out and
sends. We see different behavior between UTS and BFS.

Recall that global memory can be accessed in two ways. For requests of
locations that are expected to shared, delegate operations are
used. For requests of non-shared locations where locality can be
exploited, explicit cache operations are used. We see this reflected
in the fraction of requests that are handled remotely: less than one
percent of delegate operations in both BFS and UTS access local
data. Since explicit cache operations are intended to exploit
locality; they more often access data on the local node. Only 20
percent of UTS explicit cache operations are remote, whereas 44
percent of BFS explict cache operations are remote.

In UTS, we obtained the best throughput by flushing frequently: on the
order 100 microseconds. This, combined with the fewer number of remote
accesses, means that there are a limited nubmer of requests generated
per flush interval, leading to packets primarily smaller than 10KB.

In BFS, we see a trimodal distribution. There are two reasons for
this. First, in BFS, we obtained the best throughput by flushing at
least 500 microseconds apart, and also flushing whenever the core went
idle. Since this process ignores whether another flush has occured
recently, it leads to a large number of small messages. This leads to
the first and smallest mode of the distribution.

The other two modes are due to a locality optimization in the \Grappa
version of BFS. To minimize unnecessary remote communication, each
core visits vertices whoes edge lists are stored locally. Since we are
traversing power-law graphs, a small fraction of vertices have a very
high number of edges. This means that occasionally a core visits a
high-degree vertex which allows it to very quickly generate a large
batch of requests. This interacts well with aggregation; when we find
high degree vertices, we are able to aggregate packets on the order of
40KB. In flush intervals without high-degree vertices, we see packets
on the order of 8KB.

>>>>

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{results/benchmarks.pdf}
    \end{center}
    \caption{Comparing \Grappa with XMT, hand-tuned MPI and UPC.}
    \label{fig:grappa-comparisons}
\end{figure}

Figure~\ref{fig:grappa-comparisons} shows the results. Overall, \Grappa is
within striking distance of XMT and well beyond UPC performance (UTS only).
However, it underperforms compared to hand-tuned MPI. Nevertheless, this
comparison needs to be taken with a grain of salt for several reasons. First,
we have not spent as much time tuning \Grappa's implementations. This is
supported by inspecting the BFS MPI implementation we used, which clearly
shows that it employs a lot of algorithmic-specific optimizations. In fact,
some of these optimizations resemble some of what \Grappa does automatically,
like message aggregation. Second, the GUPS results presented earlier suggests
that \Grappa performance can do a lot better. 
% \TODO{update this once we have more results. }

@@

To characterize the effectiveness of aggregation, we collected
histograms of aggregated message sizes and application memory request
latency for BFS and UTS (Figure \ref{fig:grappa-message-size}).

%%% I couldn't make this paragraph useful in the limited time we have.
%%% I think it's ok without. -Mark
\comment{Recall that global memory can be accessed in two ways. For requests of
locations that are expected to shared, delegate operations are
used. For requests of non-shared locations where locality can be
exploited, explicit cache operations are used. We see this reflected
in the fraction of requests that are handled remotely: less than one
percent of delegate operations in both BFS and UTS access local
data. Since explicit cache operations are intended to exploit
locality; they more often access data on the local node. Only 20
percent of UTS explicit cache operations are remote, whereas 44
percent of BFS explict cache operations are remote.}

In UTS, we obtained the best throughput by flushing frequently: on the
order of 100 microseconds. This, combined with the fewer number of remote
accesses, means that there are a limited nubmer of requests generated
per flush interval, leading to packets primarily smaller than 10KB.

In BFS, we see a trimodal distribution. There are two reasons for this. First, in BFS, we obtained the best throughput by flushing at least 500 microseconds apart, and also flushing whenever a core is idle. Since this process ignores whether another flush has occurred recently, it leads to a large number of small messages. This leads to the first and smallest mode of the distribution.  The other two modes are due to a locality optimization in the \Grappa version of BFS. To minimize unnecessary remote communication, each core visits vertices whoes edge lists are stored locally. Since we are traversing power-law graphs, a small fraction of vertices have a very high number of edges. This means that occasionally a core visits a high-degree vertex which allows it to very quickly generate a large batch of requests. This interacts well with aggregation; when we find high degree vertices, we are able to aggregate packets on the order of 40KB. In flush intervals without high-degree vertices, we see packets on the order of 8KB.
